On The Efficiency of Jet Production in Radio Galaxies

### 2015-10-19 paper.tex

% Assessing the Kinetic Efficiencies of Low-luminosity AGNs: Evidence for Magnetically Arrested Accretion
% The Kinetic Efficiencies of Radio Galaxies: Evidence for Magnetically Arrested Accretion
% Kinetic Efficiencies of Radio Galaxies: Evidence for Blandford-Znajek powering of jets
\title[On the Efficiency of Jet Production in AGNs]{On The Efficiency of Jet Production in Radio Galaxies} 
\author[Nemmen]{Rodrigo S. Nemmen,$^{1,2,3,4}$\thanks{E-mail: rodrigo.nemmen@iag.usp.br} Alexander Tchekhovskoy$^{5,6,7}$ \\
$^{1}$Instituto de Astronomia, Geof\'{\i}sica e Ci\^encias Atmosf\'ericas, Universidade de S\~ao Paulo, S\~ao Paulo, SP 05508-090, Brazil \\
$^{2}$NASA Goddard Space Flight Center, Greenbelt, MD 20771, USA \\
$^{3}$Center for Research and Exploration in Space Science \& Technology (CRESST) \\
$^{4}$Department of Physics, University of Maryland, Baltimore County, 1000 Hilltop Circle, Baltimore, MD 21250, USA \\
$^5$Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA \\$^6$Department of Astronomy and Theoretical Astrophysics Center, University of California, Berkeley, CA 94720-3411, USA \\
$^7$NASA Einstein Fellow \\
}

@@

% Measuring jet efficiencies: why? importance
% inspired on bullet points of slide #7 from talk to Ramesh's group, Nov. 2013
One promising approach for probing the physical conditions behind jet launching in AGNs is to evaluate the energy efficiency of jet production which we define as $\eta_{\rm jet} \equiv P_{\rm jet}/(\dot{M}_{\rm BH} c^2)$, given appropriate measurements of the jet power $P_{\rm jet}$ \emph{and} the mass accretion rate onto the black hole $\dot{M}_{\rm BH}$. Estimating the efficiency of jet production gives us an indirect diagnostic of the jet formation process and the source of jet power. For instance, high efficiencies would suggest spin powering and possibly dynamically important magnetic fields near the horizon (e.g., \citealt{Sasha11,Zamaninasab14}) as opposed to an outflow from an accretion disk \citep{Blandford82,Ghosh97,Livio99}. $\eta_{\rm jet}$ is also relevant for our understanding of the ``radio mode'' of AGN feedback in massive early-type galaxies (e.g., \citealt{Allen06,McNamara07,Fabian12}) and it impacts the cosmological evolution and growth of massive black holes (e.g., \citealt{Fanidakis11,Volonteri12}).

@@

@@

% reason for increased scatter
There are a couple of possible explanations for the significant scatter in the $\dot{M}_B c^2 - P_{\rm jet}$ relation. There is the possibility that the scatter is primarily due to the uncertainties in the determination of the cavity power and the densities at the Bondi radius, as discussed in \cite{Russell13} (see their section 4.3). Nonetheless, it is also possible that the scatter has a physical origin and there are a couple of possible physical processes that could contribute to it:

@@

We acknowledge useful discussions with Mihoko Yukita, Ka-Wah Wong, Aleksander Sadowski, Helen Russell, Jonathan C.~McKinney, Jeremy Schnittman, Markos Georganopoulos, Ramesh Narayan, Jeff McClintock, Feng Yuan and Sylvain Guiriec. RSN was partially supported by the NASA Postdoctoral Program (NPP) at Goddard Space Flight Center, administered by Oak Ridge Associated Universities through a contract with NASA, as well as the NASA grant NNH10ZDA001N and FAPESP. AT was supported by by NASA through Einstein Postdoctoral Fellowship grant number PF3-140115 awarded by the Chandra X-ray Center, which is operated by the Smithsonian Astrophysical Observatory for NASA under contract NAS8-03060, and NASA support via High-End Computing (HEC) Program through the NASA Advanced Supercomputing (NAS) Division at Ames Research Center that provided access to the Pleiades supercomputer, as well as NSF support through an XSEDE computational time allocation TG-AST100040 on NICS Kraken, Nautilus, TACC Stampede, Maverick, and Ranch. This project made considerable use of IPython \citep{ipython} and the Astropy and mcerp libraries.

>>>>

% Assessing the Kinetic Efficiencies of Low-luminosity AGNs: Evidence for Magnetically Arrested Accretion
% The Kinetic Efficiencies of Radio Galaxies: Evidence for Magnetically Arrested Accretion
% Kinetic Efficiencies of Radio Galaxies: Evidence for Blandford-Znajek powering of jets
\title[On the Efficiency of Jet Production in AGNs]{On The Efficiency of Jet Production in Radio Galaxies} 
\author[Nemmen]{Rodrigo S. Nemmen,$^{1,2,3,4}$\thanks{E-mail: rodrigo.nemmen@iag.usp.br} Alexander Tchekhovskoy$^{5,6,7}$, 
  Henrique S. Xavier$^{1}$\\
$^{1}$Instituto de Astronomia, Geof\'{\i}sica e Ci\^encias Atmosf\'ericas, Universidade de S\~ao Paulo, S\~ao Paulo, SP 05508-090, Brazil \\
$^{2}$NASA Goddard Space Flight Center, Greenbelt, MD 20771, USA \\
$^{3}$Center for Research and Exploration in Space Science \& Technology (CRESST) \\
$^{4}$Department of Physics, University of Maryland, Baltimore County, 1000 Hilltop Circle, Baltimore, MD 21250, USA \\
$^5$Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA \\$^6$Department of Astronomy and Theoretical Astrophysics Center, University of California, Berkeley, CA 94720-3411, USA \\
$^7$NASA Einstein Fellow \\
}

@@

\subsection{New subsection}
% Measuring jet efficiencies: why? importance
% inspired on bullet points of slide #7 from talk to Ramesh's group, Nov. 2013
One promising approach for probing the physical conditions behind jet launching in AGNs is to evaluate the energy efficiency of jet production which we define as $\eta_{\rm jet} \equiv P_{\rm jet}/(\dot{M}_{\rm BH} c^2)$, given appropriate measurements of the jet power $P_{\rm jet}$ \emph{and} the mass accretion rate onto the black hole $\dot{M}_{\rm BH}$. Estimating the efficiency of jet production gives us an indirect diagnostic of the jet formation process and the source of jet power. For instance, high efficiencies would suggest spin powering and possibly dynamically important magnetic fields near the horizon (e.g., \citealt{Sasha11,Zamaninasab14}) as opposed to an outflow from an accretion disk \citep{Blandford82,Ghosh97,Livio99}. $\eta_{\rm jet}$ is also relevant for our understanding of the ``radio mode'' of AGN feedback in massive early-type galaxies (e.g., \citealt{Allen06,McNamara07,Fabian12}) and it impacts the cosmological evolution and growth of massive black holes (e.g., \citealt{Fanidakis11,Volonteri12}).

@@

\subsection{Git and github challange}

We are at IAG playing around with git and github. Today is 19/10/2015 and I am Henrique. I will also change the next section (the scatter iin eta).

@@

% reason for increased scatter
There are a couple of possible explanations for the significant scatter in the $\dot{M}_B c^2 - P_{\rm jet}$ relation. There is the possibility that the scatter is primarily due to the uncertainties in the determination of the cavity power and the densities at the Bondi radius, as discussed in \cite{Russell13} (see their section 4.3). NOTE!! This is a change due to playing with github. Nonetheless, it is also possible that the scatter has a physical origin and there are a couple of possible physical processes that could contribute to it:

@@

Eu agradeco ao meu pai e a minha mae. Um beijo no coracao de todo o mundo. Git eh muito util. etc etc
We acknowledge useful discussions with Mihoko Yukita, Ka-Wah Wong, Aleksander Sadowski, Helen Russell, Jonathan C.~McKinney, Jeremy Schnittman, Markos Georganopoulos, Ramesh Narayan, Jeff McClintock, Feng Yuan and Sylvain Guiriec (who the f&&k are these people????). RSN was partially supported by the NASA Postdoctoral Program (NPP) at Goddard Space Flight Center, administered by Oak Ridge Associated Universities through a contract with NASA, as well as the NASA grant NNH10ZDA001N and FAPESP. AT was supported by by NASA through Einstein Postdoctoral Fellowship grant number PF3-140115 awarded by the Chandra X-ray Center, which is operated by the Smithsonian Astrophysical Observatory for NASA under contract NAS8-03060, and NASA support via High-End Computing (HEC) Program through the NASA Advanced Supercomputing (NAS) Division at Ames Research Center that provided access to the Pleiades supercomputer, as well as NSF support through an XSEDE computational time allocation TG-AST100040 on NICS Kraken, Nautilus, TACC Stampede, Maverick, and Ranch. This project made considerable use of IPython \citep{ipython} and the Astropy and mcerp libraries.
